Variants on the Flight and Maintenance Planning problem

### 2018-09-12 latex/notes/model/model.tex
updated model and simulation.
updated some notes and keep advancing with iimplementation of the model and data.
new version of model with minimul consumption

    \begin{tabular}{p{8mm}p{167mm}}
        $H_j$             & amount of resource time required by task $j$. \\
        $R_j$             & number of resources required by task $j$. \\
        $MT_j$            & minimum number of consecutive periods a resource can be assigned to task $j$. \\
        % $aut_i$           & maximal absolute usage time for resource $i$. \\
        % $aet_i$           & maximal absolute elapsed time for resource $i$. \\
        $M$               & maintenance duration in number of periods. \\
        $C^{max}$         & maximum number of simultaneous maintenance operations. \\
        % $E$               & remaining elapsed time after a maintenance. \\
        $E^{M}$         & maximum number of periods between two consecutive maintenances. \\
        $E^{m}$         & minimum number of periods between two consecutive maintenances. \\
        $H^{M}$               & remaining usage time after a maintenance. \\
        % $S$               & Storage duration in number of periods. \\
        % $W_1$             & weight of the first objective in the objective function. \\
        % $W_2$             & weight of the second objective in the objective function. \\
        $N_t$             & number of resources in already-planned maintenances in period $t$.\\
        $N^{clust}_{kt}$         & number of resources in cluster $k$ in already-planned maintenances in period $t$.\\
        % $D_t$             & number of resources to be assigned in total in period $t$. \\
        $A^{clust}_{kt}$         & maximum number of resources in cluster $k$ that can be simultaneously under maintenance at period $t$.\\
        $H^{clust}_{kt}$         & minimum number of total remaining usage time for cluster $k$ at period $t$.\\
        $Rut^{Init}_{i}$  & remaining usage time for resource $i$ at the start of the planning horizon. \\
        $Ret^{Init}_{i}$  & remaining elapsed time for resource $i$ at the start of the planning horizon. \\
        % $Ret^{Init}_{sum}$& sum of remaining elapsed times at the start of the planning horizon. \\
        % $Rut^{Init}_{sum}$& sum of remaining elapsed time at the start of the planning horizon. \\
    \end{tabular}

@@

    \begin{tabular}{p{8mm}p{167mm}}
        $a_{jti}$   &  =1 if task $j \in J$ in period $t \in \mathcal{T}_j$ is realized with resource $i \in \mathcal{I}_j$, 0 otherwise. \\  
        $a^s_{jti}$ &  =1 if resource $i$ starts a new assignment to task $j$ in period $t$. If $a_{jti} = 1$ and $a_{j(t-1)i} = 0$. \\  
        $m_{it}$    &  =1 if resource $i \in I$ starts a maintenance operation in period $t \in \mathcal{T}$, 0 otherwise. \\
        % $s_{it}$    &  =1 if resource $i \in I$ starts a storage operation in period $t \in \mathcal{T}$, 0 otherwise. \\
        $rut_{it}$  &  remaining usage time (continuous) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. \\  
        % $ret_{it}$  &  remaining elapsed time (integer) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. \\  
        % $u_{max}$   &  maximal number (integer) of unavailable resources in any period. \\
        % $m_{max}$   &  maximal number (integer) of resources in maintenance in any period. \\
    \end{tabular}

@@

@@

    Constraints (\ref{eq:avalaibility-cluster}) guarantee a minimum availability of resources for each cluster $k$. A cluster is defined by the biggest group of resources that is required, exclusively, by at least one task. Constraints (\ref{eq:avalaibility-hours-cluster}) guarantee that a minimum availability in terms of usage houres for each cluster $k$.

    \begin{align}
        % remaining used time
         & rut_{it} \leq rut_{i(t-1)} + H^{M} m_{it} - \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j 
            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rut_upper}\\
        & rut_{i0} = Rut^{Init}_i
               & i \in \mathcal{I} \label{eq:rut_initial}\\
        & rut_{it} \geq H^{M} m_{it'}
                & t \in \mathcal{T}, t' \in \mathcal{T}^{s}_t, i \in \mathcal{I}\label{eq:rut_lower}\\ 
        & rut_{it} \in [0,H^{M}]
                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:mu}              
        % & ret_{it} \leq ret_{it-1} - 1 + \sum_{t' \in \mathcal{T}^{st}_t} s_{it'} + E m_{it}
        %         & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:ret_upper}\\
        % & ret_{i0} = Ret^{Init}_i
        %         & i \in \mathcal{I} \label{eq:ret_initial}\\
        % & ret_{it} \geq E m_{it}
        %         & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_lower}\\                 
        % & ret_{it} \in [0,E]
        %         & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:me}\\
        % & \sum_{i \in \mathcal{I}} ret_{it} \geq Ret^{Init}_{sum}
        %       & t = |\mathcal{T}| \label{eq:min_ret}\\
        % & \sum_{i \in \mathcal{I}} rut_{i|\mathcal{T}|} \geq Rut^{Init}_{sum}\\
    \end{align}

        % These constraints calculate the balances of hours for each resource.
    The remaining usage time is defined by (\ref{eq:rut_upper})-(\ref{eq:rut_initial}) and its limits by (\ref{eq:rut_lower})-(\ref{eq:mu}).

>>>>

    \begin{tabular}{p{8mm}p{167mm}}
        $H_j$             & amount of resource time required by task $j$. \\
        $R_j$             & number of resources required by task $j$. \\
        $MT_j$            & minimum number of consecutive periods a resource can be assigned to task $j$. \\
        $u^{min}$       & Absolut minimum usage time of resource when resources is not in maintenance.\\
        % $aut_i$           & maximal absolute usage time for resource $i$. \\
        % $aet_i$           & maximal absolute elapsed time for resource $i$. \\
        $M$               & maintenance duration in number of periods. \\
        $C^{max}$         & maximum number of simultaneous maintenance operations. \\
        % $E$               & remaining elapsed time after a maintenance. \\
        $E^{M}$         & maximum number of periods between two consecutive maintenances. \\
        $E^{m}$         & minimum number of periods between two consecutive maintenances. \\
        $H^{M}$               & remaining usage time after a maintenance. \\
        % $S$               & Storage duration in number of periods. \\
        % $W_1$             & weight of the first objective in the objective function. \\
        % $W_2$             & weight of the second objective in the objective function. \\
        $N_t$             & number of resources in already-planned maintenances in period $t$.\\
        $N^{clust}_{kt}$         & number of resources in cluster $k$ in already-planned maintenances in period $t$.\\
        % $D_t$             & number of resources to be assigned in total in period $t$. \\
        $A^{clust}_{kt}$         & maximum number of resources in cluster $k$ that can be simultaneously under maintenance at period $t$.\\
        $H^{clust}_{kt}$         & minimum number of total remaining usage time for cluster $k$ at period $t$.\\
        $Rut^{Init}_{i}$  & remaining usage time for resource $i$ at the start of the planning horizon. \\
        $Ret^{Init}_{i}$  & remaining elapsed time for resource $i$ at the start of the planning horizon. \\
        % $Ret^{Init}_{sum}$& sum of remaining elapsed times at the start of the planning horizon. \\
        % $Rut^{Init}_{sum}$& sum of remaining elapsed time at the start of the planning horizon. \\
    \end{tabular}

@@

    \begin{tabular}{p{8mm}p{167mm}}
        $a_{jti}$   &  =1 if task $j \in J$ in period $t \in \mathcal{T}_j$ is realized with resource $i \in \mathcal{I}_j$, 0 otherwise. \\  
        $a^s_{jti}$ &  =1 if resource $i$ starts a new assignment to task $j$ in period $t$. If $a_{jti} = 1$ and $a_{j(t-1)i} = 0$. \\  
        $m_{it}$    &  =1 if resource $i \in I$ starts a maintenance operation in period $t \in \mathcal{T}$, 0 otherwise. \\
        $u_{it}$    &  usage time consumed for resource $i \in I$ during period $t \in \mathcal{T}$. \\        
        % $s_{it}$    &  =1 if resource $i \in I$ starts a storage operation in period $t \in \mathcal{T}$, 0 otherwise. \\
        $rut_{it}$  &  remaining usage time (continuous) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. \\  
        % $ret_{it}$  &  remaining elapsed time (integer) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. \\  
        % $u_{max}$   &  maximal number (integer) of unavailable resources in any period. \\
        % $m_{max}$   &  maximal number (integer) of resources in maintenance in any period. \\
    \end{tabular}

@@

  % TODO: say something about fixed tasks at the beginning.
  % TODO: say something about fixed maintenances at the beginning.

@@

    Constraints (\ref{eq:avalaibility-cluster}) guarantee a minimum availability of resources for each cluster $k$. A cluster is defined by the biggest group of resources that is required, exclusively, by at least one task. Constraints (\ref{eq:avalaibility-hours-cluster}) guarantee that a minimum availability in terms of usage hours for each cluster $k$.

    \begin{align}
        % remaining used time
         & u_{it} \geq \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j 
            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:usage}\\
         & u_{it} \geq u^{min} (1 - \sum_{t' \in \mathcal{T}^{s}_t} m_{it'})
            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:usage}\\
         & u_{it} \in [0, \max_j{\{H_j\}}]
            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:usage_lower}\\
         & rut_{it} \leq rut_{i(t-1)} + H^{M} m_{it} - u_{it}
            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rut_upper}\\
        & rut_{i0} = Rut^{Init}_i
               & i \in \mathcal{I} \label{eq:rut_initial}\\
        & rut_{it} \geq H^{M} m_{it'}
                & t \in \mathcal{T}, t' \in \mathcal{T}^{s}_t, i \in \mathcal{I}\label{eq:rut_lower}\\ 
        & rut_{it} \in [0,H^{M}]
                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:mu}              
        % & ret_{it} \leq ret_{it-1} - 1 + \sum_{t' \in \mathcal{T}^{st}_t} s_{it'} + E m_{it}
        %         & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:ret_upper}\\
        % & ret_{i0} = Ret^{Init}_i
        %         & i \in \mathcal{I} \label{eq:ret_initial}\\
        % & ret_{it} \geq E m_{it}
        %         & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_lower}\\                 
        % & ret_{it} \in [0,E]
        %         & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:me}\\
        % & \sum_{i \in \mathcal{I}} ret_{it} \geq Ret^{Init}_{sum}
        %       & t = |\mathcal{T}| \label{eq:min_ret}\\
        % & \sum_{i \in \mathcal{I}} rut_{i|\mathcal{T}|} \geq Rut^{Init}_{sum}\\
    \end{align}

        % These constraints calculate the balances of hours for each resource.
    The usage time per resource and period is calculated in (\ref{eq:usage})-(\ref{eq:usage_lower}). The remaining usage time is defined by (\ref{eq:rut_upper})-(\ref{eq:rut_initial}) and its limits by (\ref{eq:rut_lower})-(\ref{eq:mu}).
